---
title: "Analyze Results"
author: "Basketball Salaries Team"
output: pdf_document
geometry: margin=.5cm
---

## Load Results from Optimal ElasticNet Model
```{r}
df_enet <- read.csv('../data/predictions/elasticnet.csv')
head(df_enet)
```

## Plot Salary vs Predicted Salary
```{r message=F}
library(ggplot2)
ggplot(df_enet,aes(x=salary,y=salary_hat_elasticnet)) + 
  geom_point() +
  geom_smooth()
ggsave("../figures/elasticnet_salary_vs_prediction_scatter.png", width=10, height=7)
```

## Plot Salary vs Redisual Prediction
Appears to be a positive linear trend between the salary and residuals. \
Thus, the optimal elasticnet model tends to underestimate players with higher salaries.  
```{r message=F}
library(ggplot2)
residual <- df_enet$salary - df_enet$salary_hat_elasticnet
ggplot(df_enet, aes(x=salary,y=residual)) + 
  geom_point() + 
  geom_smooth()
ggsave("../figures/elasticnet_salary_vs_residual_scatter.png", width=10, height=7)
```

## Underrated Players in 2016
An underrated player is any player who should be making 50% more\
$\hat{salary}-salary > \frac{1}{2} salary$\
Players who our model classifies as underrated tend to get higher salaries the next year
```{r}
# get underrated players in 2016
df_underrated_2016 <- df_enet[
  ((df_enet$salary_hat_elasticnet-df_enet$salary) > (df_enet$salary/2)) & 
  df_enet$year==2016,]
# link with 2017 stats
df_underrated_2017 <- df_enet[(df_enet$name%in%df_underrated_2016$name)&(df_enet$year==2017),]
df_underrated <- merge(df_underrated_2016,df_underrated_2017,all.x=F,all.y=F,by='name')
names(df_underrated) <- c('name','2016','salary_2016','salary_hat_2016','2017','salary_2017','salary_hat_2017')
df_underrated <- df_underrated[,c('name','salary_2016','salary_hat_2016','salary_2017','salary_hat_2017')]
df_underrated
# calculate percent increase factor
sal_factor_2016 = .25
p_increase <- 100*mean(
  (df_underrated$salary_2017-df_underrated$salary_2016) > sal_factor_2016*df_underrated$salary_2016)
sprintf(
  'Percent of underrated players in 2016 whose salary increased by more than 25%% for 2017: %.1f%%',p_increase)
```


## Overrated Players in 2016
An overrated player is any player who should be making 50% more\
$salary-\hat{salary} > \frac{1}{2} salary$\
Hard for players who have high salaries to get a salary cut despite lacking stats
```{r}
df_overrated_216 <- df_enet[
  ((df_enet$salary-df_enet$salary_hat_elasticnet) > (df_enet$salary/2)) & 
  df_enet$year==2016,]
# link with 2017 stats
df_overrated_2017 <- df_enet[(df_enet$name%in%df_overrated_216$name)&(df_enet$year==2017),]
df_overrated <- merge(df_overrated_216,df_overrated_2017,all.x=F,all.y=F,by='name')
names(df_overrated) <- c('name','2016','salary_2016','salary_hat_2016','2017','salary_2017','salary_hat_2017')
df_overrated <- df_overrated[,c('name','salary_2016','salary_hat_2016','salary_2017','salary_hat_2017')]
df_overrated
# calculate percent decrease factor
p_decrease <- 100*mean((df_overrated$salary_2016-df_overrated$salary_2017) > 0)
sprintf('Percent of overrated players in 2016 whose salary decreased for 2017: %.1f%%',p_decrease)
```