---
title: "Models and Deployment"
author: "Basketball Salaries Team"
output: pdf_document
geometry: margin=.5cm
---

## Load Primary Dataset
```{r}
df_p.all <- read.csv('../data/pooled/primary.csv')
df_p.all$year <- as.factor(df_p.all$year)
str(df_p.all)
head(df_p.all)
```

## Load Complete (Primary + Secondary) Dataset
```{r}
df_c.all <- read.csv('../data/pooled/complete.csv')
df_c.all$year <- as.factor(df_c.all$year)
str(df_c.all)
head(df_c.all)
```

## Split Primary & Complete Datasets into Train Test
```{r message=F}
library(caret)
set.seed(7)
# primary dataset
train_rows.p <- createDataPartition(y=df_p.all[,'salary'], list=FALSE, p=.8)
df_p.train <- df_p.all[train_rows.p,]
df_p.test <- df_p.all[-train_rows.p,]
nrow(df_p.all)
nrow(df_p.train)
nrow(df_p.test)
# complete dataset
train_rows.c <- createDataPartition(y=df_c.all[,'salary'], list=FALSE, p=.8)
df_c.train <- df_c.all[train_rows.c,]
df_c.test <- df_c.all[-train_rows.c,]
nrow(df_c.all)
nrow(df_c.train)
nrow(df_c.test)
```

## Load Train/Test Datasets Resulting from Variable Selection
```{r}
df_p_vs.train = read.csv('../data/train_test/primary/train.csv')
df_p_vs.train$year <- as.factor(df_p_vs.train$year)
df_p_vs.test = read.csv('../data/train_test/primary/test.csv')
df_p_vs.test$year <- as.factor(df_p_vs.test$year)
df_c_vs.train = read.csv('../data/train_test/complete/train.csv')
df_c_vs.train$year <- as.factor(df_c_vs.train$year)
df_c_vs.test = read.csv('../data/train_test/complete/test.csv')
df_c_vs.test$year <- as.factor(df_c_vs.test$year)
```

### Modeling Helper Functions
```{r}
r_squared <- function(y,yHat){1-sum((y-yHat)^2)/sum((y-mean(y))^2)}
mse <- function(y,yHat){mean((y-yHat)^2)}
model_results <- function(model,dataset,y,yHat){
  r2_test <- r_squared(y,yHat)
  mse_test <- mse(y,yHat)
  cat(sprintf('Model: %-25s Dataset: %-15s R^2 Test: %-10.3f MSE: %-10.3e\n',model,dataset,r2_test,mse_test))}
```

## Simple Linear Regression Models
```{r}
# modeling function
slr_modeling <- function(dataset,df_train,df_test){
  model <- 'SLR'
  x_vars <- names(df_train)[!(names(df_train)%in%c('name','salary','X2P','X2PA','TRB','PTS'))]
  f <- as.formula(sprintf('salary ~ `%s`',paste(x_vars,collapse='` + `')))
  slr_model <- lm(f,data=df_train)
  yhat <- predict(slr_model,df_test)
  model_results(model,dataset,df_test[['salary']],yhat)
  return(slr_model)}
# train/test Simple Linear Regression models
names(df_p.train)
ignore <- slr_modeling('primary',df_p.train,df_p.test)
ignore <- slr_modeling('complete',df_c.train,df_c.test)
names(df_p_vs.train)
ignore <- slr_modeling('primary VS',df_p_vs.train,df_p_vs.test) # primary variable subset
ignore <- slr_modeling('complete VS',df_c_vs.train,df_c_vs.test) # complete variable subset
```

## Lasso, Ridge, and Elastic Net Models with 10-fold Cross validation for alpha = seq(0,1,by=.05)
```{r message=F}
library(glmnet)
# modeling function
lre_modeling <- function(dataset,x_train,y_train,x_test,y_test,alphas,mkplot){
  # fit models
  for (i in alphas){
    set.seed(7) # seed for reproducibility
    model_name <- sprintf('fit_alpha_%.2f',i)
    assign(model_name, cv.glmnet(x_train, y_train, type.measure="mse",alpha=i,family="gaussian"))
    model <- get(model_name)
    yhat <- predict(model,s=model$lambda.min,newx=x_test)
    model_results(model_name,dataset,y_test,yhat)
    # plot
    if(mkplot){
      path = sprintf("../figures/elasticnet_models/alpha_%.2f.%s.png",i,dataset)
      png(file=path)
      par(mfrow=c(2,1))
      glmnet_model <- glmnet(x_train, y_train, family="gaussian",alpha=i)
      plot(glmnet_model)
      title(sprintf('Elasticnet Model, %s dataset, alpha = %.2f',dataset,i),line=3)
      plot(model,xvar='lambda')      
      dev.off()}}
  return(model)} # return final model created
# extract train/test datasets of only numeric variables as required by glmnet models
#    primary dataset
numeric_vars.p <- names(Filter(is.numeric,df_p.train))
numeric_x_vars.p <- numeric_vars.p[!(numeric_vars.p%in%c('salary'))]
x_train.p <- data.matrix(df_p.train[,numeric_x_vars.p])
y_train.p <- df_p.train[['salary']]
x_test.p <- data.matrix(df_p.test[,numeric_x_vars.p])
y_test.p <- df_p.test[['salary']]
#    complete dataset
numeric_vars.c <- names(Filter(is.numeric,df_c.train))
numeric_x_vars.c <- numeric_vars.c[!(numeric_vars.c%in%c('salary'))]
x_train.c <- data.matrix(df_c.train[,numeric_x_vars.c])
y_train.c <- df_c.train[['salary']]
x_test.c <- data.matrix(df_c.test[,numeric_x_vars.c])
y_test.c <- df_c.test[['salary']]
# train/test Simple Linear Regression models
mkplots <- FALSE # change to TRUE if you want to generate plots
ignore <- lre_modeling('primary',x_train.p,y_train.p,x_test.p,y_test.p,seq(0,1,by=.05),mkplots)
ignore <- lre_modeling('complete',x_train.c,y_train.c,x_test.c,y_test.c,seq(0,1,by=.05),mkplots)
```

## Save Optimal Model to File
Optimal model with largest R^2 Test and smallest MSE is SLR on Complete Backwards Selected Variables
```{r}
optimal_model <- slr_modeling('complete VS',df_c_vs.train,df_c_vs.test) # complete variable subset
str(df_c_vs.train)
saveRDS(optimal_model,file='../data/optimal_model/slr/model.rds')
# save factor levels for input encoding 
saveRDS(levels(df_c_vs.train$year),file='../data/optimal_model/slr/year_levels.rds')
saveRDS(levels(df_c_vs.train$Pos),file='../data/optimal_model/slr/Pos_levels.rds')
saveRDS(levels(df_c_vs.train$Tm),file='../data/optimal_model/slr/Tm_levels.rds')
```

## Deployment of Optimal Model
```{r eval=F}
library(plumber)
r <- plumb("./deploy_optimal_model.R")
r$run(port=8000)
```

### Single Player Prediction Command
```{r eval=F}
# 2016 LeBron James Example
df_c_vs.train[df_c_vs.train$name=='lebron james'&df_c_vs.train$year=='2016',]
cmd <- "curl -X POST 'http://localhost:8000/predict_salary?\
        year=2016&POS=SF&Age=31&Tm=CLE&G=76&MP=2709X3PAr=0.199&\
        TOV.=13.2&DWS=4&DBPM=2.3&BPM=9.1&FG=737&STL=104&PF=143&ovr=99'"
```

## Prediction from Optimal Model
```{r}
df_c.all.copy <- df_c.all
df_c.all.copy$salary_hat <- predict(optimal_model,df_c.all.copy)
df_c.all.copy <- df_c.all.copy[,c('name','year','salary','salary_hat')]
head(df_c.all.copy)
write.csv(df_c.all.copy,'../data/predictions/slr.csv',row.names=F)
```

